---
title: "1DTT"
output:
  word_document: default
  pdf_document: default
date: "2024-03-09"
---

1) (Modified Exercise 14.4 in ESL) Cluster the demographic data
(>data(marketing in ESL package)) of Table 14.1 using “generalized
association rules” with a classification tree. (This data can also be found on
the ESL website). Specifically, generate a reference sample the same size as
the training set, by either (a) randomly permuting the columns independently,
or by (b) sampling from a uniform distribution the values within each feature.
Build a classification tree to the training sample (class 1) and the reference
sample (class 0) and describe the terminal nodes having highest estimated
class 1 probability?



```{r}


```



```{r}
rm(list = ls())

library(rpart) 
library(rpart.plot)
```


```{r}

col= c("ANNUAL_INCOME", "Sex", "Marital_Status", "Age", "Education", "Occupation", "Years_In_BayArea", "Dual_Income", "Numbers_in_Household",
                  "Number_of_Children", "Householder_Status", "Type_of_Home", "Ethnic_Classification", "Language_in_Home")

data = read.table("/Users/avishma/Downloads/marketingdata", header=FALSE, sep=" ", na.strings = "NA", col.names = col)
head(data)


str(data)


```


##Data Preprocessing


```{r}

any_na = any(is.na(data))
na_count_per_column <- colSums(is.na(data))

any_na
na_count_per_column

```



```{r}

# Treating numerical variables (median)

data$ANNUAL_INCOME[is.na(data$ANNUAL_INCOME)] = mean(data$ANNUAL_INCOME, na.rm = TRUE)
data$Age[is.na(data$Age)] = mean(data$Age, na.rm = TRUE)
data$Years_In_BayArea[is.na(data$Years_In_BayArea)] = mean(data$Years_In_BayArea, na.rm = TRUE)
data$Numbers_in_Household[is.na(data$Numbers_in_Household)] = mean(data$Numbers_in_Household, na.rm = TRUE)


# Treating categorical variables (mode)

cat_vars = c("Sex", "Marital_Status", "Education", "Occupation", "Householder_Status", "Type_of_Home", "Ethnic_Classification", "Language_in_Home")
for (var in cat_vars) {
  data[[var]][is.na(data[[var]])] <- names(sort(table(data[[var]]), decreasing = TRUE))[1]  # Mode
}


```



```{r}

na_count <- colSums(is.na(data))
na_count

```
```{r}
str(data)
```



```{r}
categorical_vars = c("Sex", "Marital_Status", "Education", "Occupation", "Householder_Status", "Type_of_Home", "Ethnic_Classification", "Language_in_Home", "Dual_Income", "Number_of_Children")

for (var in categorical_vars) {
  data[[var]] <- as.factor(data[[var]])
}
categorical_vars
```




```{r}
set.seed(123) 
reference_sample = data.frame(sapply(data, function(column) sample(column)))
```

```{r}
data$class = 1  
reference_sample$class = 0  

combined_data = rbind(data, reference_sample)

```
```{r}
categorical_vars = c("Sex", "Marital_Status", "Education", "Occupation", "Dual_Income", "Householder_Status", "Type_of_Home", "Ethnic_Classification", "Language_in_Home")

for (var in categorical_vars) {
  combined_data[[var]] <- factor(combined_data[[var]])
}
categorical_vars 
str(categorical_vars)
```
```{r}
str(combined_data)
```

```{r}
#Categorical to factor
independent_vars <- setdiff(names(combined_data), "class")
for (var in independent_vars) {
  combined_data[[var]] <- as.factor(combined_data[[var]])
}

```

```{r}
str(combined_data)
```




#Getting the association rules:




```{r}

library(arules)

transact_list <- apply(combined_data, 1, function(row) {
  items <- names(which(row == 1)) 
  if (length(items) > 0) items else NULL
})
transact_list = Filter(Negate(is.null), transact_list)
transact = as(transact_list, "transactions")

# Apriori algorithm 
supp_rules <- apriori(transact, parameter = list(supp = 0.001, conf = 0.01))
inspect(head(sort(supp_rules, by = "support"), 4))

```





Comment on the Association rules:

Rules seem to suggest that Language_in_Home, Dual_Income, and Type_of_Home are very common attributes in the dataset, with the first one being present in almost all transactions. The presence of rules with empty antecedents (left-hand sides) could indicate that the support threshold might be set too low to capture meaningful associations, as these "rules" are simply stating the overall frequency of certain items rather than indicating a strong association between specific items.


```{r}
#Tree Model
tree_model = rpart(class ~ ., data = combined_data, method = "class")
tree_model

```



```{r}
# Plot the classification tree
rpart.plot(tree_model, main="Classification Tree", under=TRUE, faclen=0)
```

#Calculation of Class 1 Probability
```{r}
Class = function(number_of_children, numbers_in_household, householder_status, marital_status) {
  if (number_of_children == 0) {
    if (numbers_in_household %in% c(1, 2, 3, 4)) { 
      if (householder_status %in% c(1, 2)) {
        if (marital_status %in% c(2, 3, 4, 5)) {
          return(list(class = 1, probability = 0.75)) 
        }
      }
    }
  }
  
  return(list(class = 0, probability = NA)) 
}

Class(0, 2, 1, 3)

```


```{r}

summary(tree_model)

```










